\vspace{-1.5em} \bf A general and efficient representation of ancestral recombination graphs

### 2023-10-10 paper.tex
Tweak
Fig 6 fixed

Genome ARGs, on the hand,
can be either retrospective or \emph{prospective}.
The edges in a gARG represent a path of genetic inheritance from
ancestor to descendant through some
% Wilder: finds the cell-to-cell confusing. Maybe drop?
number of generations (ultimately from cell to cell),
and are annotated
with the genome coordinates over which inheritance occurs.
In a prospective gARG, the inheritance intervals $I$
on an edge joining ancestor genome $v$ and descendant $u$
defines \emph{all} regions of genome that $u$ inherited from $v$.
Thus, these  inheritance intervals are
concerned only with the genetic relationship
between the nodes in question, and are not defined with respect
to any given set of samples. We are simply recording the
genetic inheritance between two genomes
(in the simplest case, between a parent and a child).

@@

Common ancestor events occur at rate $\binom{|L|}{2}$.
In a common ancestor event, two uniformly sampled lineages have their segments
merged into a single ancestor lineage, which is added to $L$.
If the lineages have overlapping intervals of ancestry,
say, $(\ell, r, a_1)$ and $(\ell, r, a_2)$, a
\emph{coalescence} occurs. The result is a segment
$(\ell, r, a_1 + a_2)$, and if $a_1 + a_2 < n$ it is included in the
ancestor lineage. Otherwise, if $a_1 + a_2 = n$, we have found
the most recent common ancestor of all samples in the interval$[\ell, r)$
and do not need to simulate its history any further.
Non-overlapping intervals from the two lineages are included
 in the ancestor lineage without changes. Eventually,
we find resultant lineages in which all segments have fully coalesced,
and so the number of extant lineages gradually falls to zero.

@@

Figure \ref{hudson_vs_bigARG} illustrates the more complex state space
of the little ARG process, as well as the extra events which occur in the big ARG process.
Moreover, it depicts the rates of common ancestors and recombination events in each
interval of time of the realisations.
In order to evaluate these rates, and hence the sampling probability
(see e.g.\ \cite[Equation (3)]{mahmoudi2022bayesian}),
it is necessary to know the number of lineages and number of extant links
available for recombination in each time interval.
Some representations may not provide this information.
For example, in the gARG encoding depicted in Figure \ref{fig-ancestry-resolution}C
it is clear that a recombination takes place between nodes \textsf{b} and
\textsf{d} as well as \textsf{e}, but the exact time of the event is ambiguous,
and thus so is the number of lineages during the time interval. In fact, this
information cannot be recovered from the gARG encoding used in Figure \ref{fig-ancestry-resolution}C,
and alternative conventions must be introduced to resolve such ambiguities.
For the likelihood-based inference algorithms for the coalescent with recombination,
it is sufficient to create \emph{two} gARG nodes at the time of the recombination event:
see the appendix of \cite{baumdicker2021efficient} for details of evaluating
coalescent with recombination likelihoods using this convention.
This is also the interpretation depicted in
Figure \ref{hudson_vs_bigARG}, but it means that the two edges above node \textsf{b}
in Figure \ref{fig-ancestry-resolution} should correspond to only one lineage,
along which all $m-1$ links are available for recombination.
The lineage then splits into two at the time of nodes \textsf{d} and \textsf{e}.
Nodes \textsf{k}, \textsf{l}, and \textsf{m} in Figure \ref{fig-ancestry-resolution}
demonstrate that the same issue can affect the number of links available for recombination:
without an external convention, the exact time at which the trapped ancestral material on
node \textsf{k} ceases to be available for an effective recombination in the little ARG process.

@@

	\draw (0, 0) -- (0, 1.7) -- (-1,1.7) -- (1,1.7);
	\draw (-1.4, 1.5) -- (-1.1, 1.5) -- (-1.1, 1.4) -- (-1.4, 1.4) -- (-1.4, 1.5);
	\draw (-1.1, 1.45) -- (-0.6, 1.45);
	\draw (0.6, 1.45) -- (0.9, 1.45);
	\draw (0.9, 1.5) -- (1.4, 1.5) -- (1.4, 1.4) -- (0.9, 1.4) -- (0.9, 1.5);
	\draw (1.2, 1.6) -- (1.2, 1.3);
	\node [label=below:{\small $x$}] at (1.2, 1.5) {};

@@

	\draw (1, 1.7) -- (1, 3.1) -- (0,3.1) -- (2,3.1);
	\draw (-0.4, 2.85) -- (0.2, 2.85);
	\draw (0.2, 2.9) -- (0.4, 2.9) -- (0.4, 2.8) -- (0.2, 2.8) -- (0.2, 2.9);
	\draw (1.6, 2.85) -- (1.9, 2.85);
	\draw (1.9, 2.9) -- (2.2, 2.9) -- (2.2, 2.8) -- (1.9, 2.8) -- (1.9, 2.9);
	\draw (2.2, 2.85) -- (2.4, 2.85);

	\draw (-1, 1.7) -- (-1, 4.5) -- (0,4.5) -- (0,3.1);
	\node [scale=0.2,label=above left:{\small \textsf{a}}] at (-0.5,4.5) {a};
	\draw (-0.9, 4.3) -- (-0.6, 4.3) -- (-0.6, 4.2) -- (-0.9, 4.2) -- (-0.9, 4.3);
	\draw (-0.6, 4.25) -- (-0.3, 4.25);
	\draw (-0.3, 4.3) -- (-0.1, 4.3) -- (-0.1, 4.2) -- (-0.3, 4.2) -- (-0.3, 4.3);
	\draw (-0.5, 4.4) -- (-0.5, 4.1);
	\node [label=below:{\small $y$}] at (-0.5, 4.3) {};

@@

	\draw (-0.5, 4.5) -- (-0.5, 6) -- (-1.5,6) -- (0.5,6);
	\draw (-1.9, 5.75) -- (-1.3, 5.75);
	\draw (-1.3, 5.8) -- (-1.1, 5.8) -- (-1.1, 5.7) -- (-1.3, 5.7) -- (-1.3, 5.8);
	\draw (0.1, 5.8) -- (0.4, 5.8) -- (0.4, 5.7) -- (0.1, 5.7) -- (0.1, 5.8);
	\draw (0.4, 5.75) -- (0.9, 5.75);

@@

	% Edge annotations above each event
	\node [label=left:{\small$\{(0,m,1)\}$}] at (0.2,0.6) {};
	\node [label=left:{\small$\{(0,m,1)\}$}] at (4.2,0.4) {};
	\node [label=right:{\small$\{(0,m,1)\}$}] at (5.8,0.4) {};
	\node [label=left:{\small$\{(0,m,2)\}$}] at (5.2,1.6) {};
	\node [label=left:{\small$\{(0,v,1)\}$}] at (-0.8,2.3) {};
	\node [label=right:{\small$\{(v,m,1)\}$}] at (0.8,2.1) {};
	\node [label=right:{\small$\{(0,w,2)\}$}] at (3.8,3.9) {};
	\node [label=left:{\small$\{(w,m,2)\}$}] at (6.2,5) {};
	\node [label=right:{\small$\{(x,m,1)\}$}] at (-0.2,3.8) {};
	\node [label=right:{\small$\{(v,x,1)\}$}] at (1.8,4.1) {};
	\node [label=left:{\small$\{(0,v,1), (x,m,1)\}$}] at (-0.3,5.2) {};
	\node [label=right:{\small$\{(0,v,2)\}$}] at (2.8,6.4) {};
	\node [label=left:{\small$\{(x,m,1)\}$}] at (-1.3,7) {};
	\node [label=right:{\small$\{(0,v,1)\}$}] at (0.3,6.4) {};

@@

	% Numbers of extant ancestors and links, from top to bottom
	\node[label=right:{CA \; RE}] at (7.5, 8.2) {};
	\node[label=right:{$\binom{2}{2}$ \; $\frac{2 m - x - w - 2}{m - 1} \rho$}] at (7.5, 7.6) {};
	\node[label=right:{$\binom{4}{2}$ \; $\frac{2 m + 2 v - x - w - 3}{m - 1} \rho$}] at (7.5, 6.6) {};
	\node[label=right:{$\binom{3}{2}$ \; $\frac{2 m + v - w - 3}{m - 1} \rho$}] at (7.5, 5.7) {};
	\node[label=right:{$\binom{4}{2}$ \; $\frac{2 m + x - v - 4}{m - 1} \rho$}] at (7.5, 4.95) {};
	\node[label=right:{$\binom{5}{2}$ \; $\frac{2 m - 5}{m - 1} \rho$}] at (7.5, 3.8) {};
	\node[label=right:{$\binom{4}{2}$ \; $\frac{2 m - 4}{m -1} \rho$}] at (7.5, 2.75) {};
	\node[label=right:{$\binom{3}{2}$ \; $\frac{2 m - 3}{m - 1} \rho$}] at (7.5, 2.05) {};
	\node[label=right:{$\binom{2}{2}$ \; $2 \rho$}] at (7.5, 1.45) {};
	\node[label=right:{$\binom{3}{2}$ \; $3 \rho$}] at (7.5, 0.6) {};

@@

	\draw (0, 0) -- (0, 1.7) -- (-1,1.7) -- (1,1.7);
	\draw (-1.4, 1.5) -- (-1.1, 1.5) -- (-1.1, 1.4) -- (-1.4, 1.4) -- (-1.4, 1.5);
	\draw (-1.1, 1.45) -- (-0.6, 1.45);
	\draw (0.6, 1.45) -- (0.9, 1.45);
	\draw (0.9, 1.5) -- (1.4, 1.5) -- (1.4, 1.4) -- (0.9, 1.4) -- (0.9, 1.5);
	\draw (1.2, 1.6) -- (1.2, 1.3);
	\node [label=below:{\small $x$}] at (1.2, 1.5) {};

	\draw (5, 1.2) -- (5, 2.4) -- (4,2.4) -- (6,2.4);
	\draw (3.6, 2.2) -- (4.1, 2.2) -- (4.1, 2.1) -- (3.6, 2.1) -- (3.6, 2.2);
	\draw (4.1, 2.15) -- (4.4, 2.15);
	\draw (5.6, 2.15) -- (6.1, 2.15);
	\draw (6.1, 2.2) -- (6.4, 2.2) -- (6.4, 2.1) -- (6.1, 2.1) -- (6.1, 2.2);
	\draw [color=red](5.8, 2.3) -- (5.8, 2.0);
	\node [label={[red]below:{\small $z$}}] at (5.8, 2.1) {};

	\draw (1, 1.7) -- (1, 3.1) -- (0,3.1) -- (2,3.1);
	\draw (-0.4, 2.85) -- (0.2, 2.85);
	\draw (0.2, 2.9) -- (0.4, 2.9) -- (0.4, 2.8) -- (0.2, 2.8) -- (0.2, 2.9);
	\draw (1.6, 2.85) -- (1.9, 2.85);
	\draw (1.9, 2.9) -- (2.2, 2.9) -- (2.2, 2.8) -- (1.9, 2.8) -- (1.9, 2.9);
	\draw (2.2, 2.85) -- (2.4, 2.85);

	\draw (-1, 1.7) -- (-1, 4.5) -- (0,4.5) -- (0,3.1);
	\draw (-0.9, 4.3) -- (-0.6, 4.3) -- (-0.6, 4.2) -- (-0.9, 4.2) -- (-0.9, 4.3);
	\draw (-0.6, 4.25) -- (-0.3, 4.25);
	\draw (-0.3, 4.3) -- (-0.1, 4.3) -- (-0.1, 4.2) -- (-0.3, 4.2) -- (-0.3, 4.3);
	\draw (-0.5, 4.4) -- (-0.5, 4.1);
	\node [label=below:{\small $y$}] at (-0.5, 4.3) {};

@@

	% Gray dashed lines to visually separate holding times
	\draw[color=gray, dashed] (7.5, 1.2) -- (9.1, 1.2);
	\draw[color=gray, dashed] (7.5, 1.7) -- (9.1, 1.7);
	\draw[color=gray, dashed] (7.5, 2.4) -- (9.1, 2.4);
	\draw[color=gray, dashed] (7.5, 3.1) -- (9.1, 3.1);
	\draw[color=gray, dashed] (7.5, 3.8) -- (9.1, 3.8);
	\draw[color=gray, dashed] (7.5, 4.5) -- (9.1, 4.5);
	\draw[color=gray, dashed] (7.5, 5.4) -- (9.1, 5.4);
	\draw[color=gray, dashed] (7.5, 6) -- (9.1, 6);
	\draw[color=gray, dashed] (7.5, 6.6) -- (9.1, 6.6);
	\draw[color=gray, dashed] (7.5, 7.2) -- (9.1, 7.2);
\end{tikzpicture}
\end{tabular}
}
\caption{(A)
A realisation of the graph traversed by Hudson's algorithm started from a
sample of three chromosomes of length $m$ at time $t = 0$, and
propagated until time $T$. The MRCA on the genetic interval $[v, w)$ is reached
at event \textsf{b}, while that on $[0, v)$ is reached at event \textsf{c}.
The non-ancestral segment $[v, w)$ above
A contributes to the rate of effective recombinations because it
is trapped between ancestral segments. The two columns titled CA and RE
are the respective rates of mergers and recombinations when
the recombination rate is $\rho$.
(B) A corresponding realisation of a big ARG, which augments Hudson's algorithm
by tracking nonancestral lineages. The result is a simpler state space and
dynamics, at the cost of extra nodes and edges, highlighted in red, which do
not affect the local tree at any site.}
\label{hudson_vs_bigARG}
\end{figure}

>>>>

There is no particular reason, however, that ARGs must be 
retrospective. They can also be \emph{prospective}.
The edges in a gARG represent a path of genetic inheritance from
ancestor to descendant through some
% Wilder: finds the cell-to-cell confusing. Maybe drop?
number of generations (ultimately from cell to cell),
and are annotated
with the genome coordinates over which inheritance occurs.
In a prospective gARG, the inheritance intervals $I$
on an edge joining ancestor genome $v$ and descendant $u$
defines \emph{all} regions of genome that $u$ inherited from $v$.
Thus, these  inheritance intervals are
concerned only with the genetic relationship
between the nodes in question, and are not defined with respect
to any given set of samples. We are simply recording the
genetic inheritance between two genomes
(in the simplest case, between a parent and a child).

@@

Common ancestor events occur at rate $\binom{|L|}{2}$.
In a common ancestor event, two uniformly sampled lineages have their segments
merged into a single ancestor lineage, which is added to $L$.
If the lineages have overlapping intervals of ancestry,
say, $(\ell, r, a_1)$ and $(\ell, r, a_2)$, a
\emph{coalescence} occurs. The result is a segment
$(\ell, r, a_1 + a_2)$, and if $a_1 + a_2 < n$ it is included in the
ancestor lineage. Otherwise, if $a_1 + a_2 = n$, we have found
the most recent common ancestor of all samples in the interval $[\ell, r)$
and do not need to simulate its history any further.
Non-overlapping intervals from the two lineages are included
 in the ancestor lineage without changes. Eventually,
we find resultant lineages in which all segments have fully coalesced,
and so the number of extant lineages gradually falls to zero.

@@

Figure \ref{hudson_vs_bigARG} illustrates the more complex state space
of the little ARG process, as well as the extra events which occur in the big ARG process.
Moreover, it depicts the rates of common ancestors and recombination events in each
interval of time of the realisations.
In order to evaluate these rates, and hence the sampling probability
\cite[see e.g.][Equation (3)]{mahmoudi2022bayesian},
it is necessary to know the number of lineages and number of extant links
available for recombination in each time interval.
Some representations may not provide this information.
For example, in the gARG encoding depicted in Figure \ref{fig-ancestry-resolution}C
it is clear that a recombination takes place between nodes \textsf{b} and
\textsf{d} as well as \textsf{e}, but the exact time of the event is ambiguous,
and thus so is the number of lineages during the time interval. In fact, this
information cannot be recovered from the gARG encoding used in Figure \ref{fig-ancestry-resolution}C,
and alternative conventions must be introduced to resolve such ambiguities.
For the likelihood-based inference algorithms for the coalescent with recombination,
it is sufficient to create \emph{two} gARG nodes at the time of the recombination event:
see the appendix of \cite{baumdicker2021efficient} for details of evaluating
coalescent with recombination likelihoods using this convention.
This is also the interpretation depicted in
Figure \ref{hudson_vs_bigARG}, but it means that the two edges above node \textsf{b}
in Figure \ref{fig-ancestry-resolution} should correspond to only one lineage,
along which all $m-1$ links are available for recombination.
The lineage then splits into two at the time of nodes \textsf{d} and \textsf{e}.
Nodes \textsf{k}, \textsf{l}, and \textsf{m} in Figure \ref{fig-ancestry-resolution}
demonstrate that the same issue can affect the number of links available for recombination:
without an external convention, the exact time at which the trapped ancestral material on
node \textsf{k} ceases to be available for an effective recombination in the little ARG process.

@@

	\draw (0, 0) -- (0, 1.7) -- (-1,1.7) -- (1,1.7);
	\draw (-1.4, 1.5) -- (-1.1, 1.5) -- (-1.1, 1.4) -- (-1.4, 1.4) -- (-1.4, 1.5);
	\draw (-1.1, 1.45) -- (-0.6, 1.45);
	\draw (0.6, 1.45) -- (0.9, 1.45);
	\draw (0.9, 1.5) -- (1.4, 1.5) -- (1.4, 1.4) -- (0.9, 1.4) -- (0.9, 1.5);
	\draw (1.1, 1.6) -- (1.1, 1.3);
	\node [label=below:{\small $w$}] at (1.1, 1.5) {};

@@

	\draw (1, 1.7) -- (1, 3.1) -- (0,3.1) -- (2,3.1);
	\draw (-0.4, 2.85) -- (0.1, 2.85);
	\draw (0.1, 2.9) -- (0.4, 2.9) -- (0.4, 2.8) -- (0.1, 2.8) -- (0.1, 2.9);
	\draw (1.6, 2.85) -- (1.9, 2.85);
	\draw (1.9, 2.9) -- (2.1, 2.9) -- (2.1, 2.8) -- (1.9, 2.8) -- (1.9, 2.9);
	\draw (2.1, 2.85) -- (2.4, 2.85);

	\draw (-1, 1.7) -- (-1, 4.5) -- (0,4.5) -- (0,3.1);
	\node [scale=0.2,label=above left:{\small \textsf{a}}] at (-0.5,4.5) {a};
	\draw (-0.9, 4.3) -- (-0.6, 4.3) -- (-0.6, 4.2) -- (-0.9, 4.2) -- (-0.9, 4.3);
	\draw (-0.6, 4.25) -- (-0.4, 4.25);
	\draw (-0.4, 4.3) -- (-0.1, 4.3) -- (-0.1, 4.2) -- (-0.4, 4.2) -- (-0.4, 4.3);
	\draw (-0.5, 4.4) -- (-0.5, 4.1);
	\node [label=below:{\small $x$}] at (-0.5, 4.3) {};

@@

	\draw (-0.5, 4.5) -- (-0.5, 6) -- (-1.5,6) -- (0.5,6);
	\draw (-1.9, 5.75) -- (-1.4, 5.75);
	\draw (-1.4, 5.8) -- (-1.1, 5.8) -- (-1.1, 5.7) -- (-1.4, 5.7) -- (-1.4, 5.8);
	\draw (0.1, 5.8) -- (0.4, 5.8) -- (0.4, 5.7) -- (0.1, 5.7) -- (0.1, 5.8);
	\draw (0.4, 5.75) -- (0.9, 5.75);

@@

	% Edge annotations above each event
	\node [label=left:{\small$\{(0,m,1)\}$}] at (0.2,0.6) {};
	\node [label=left:{\small$\{(0,m,1)\}$}] at (4.2,0.4) {};
	\node [label=right:{\small$\{(0,m,1)\}$}] at (5.8,0.4) {};
	\node [label=left:{\small$\{(0,m,2)\}$}] at (5.2,1.6) {};
	\node [label=left:{\small$\{(0,v,1)\}$}] at (-0.8,2.3) {};
	\node [label=right:{\small$\{(v,m,1)\}$}] at (0.8,2.1) {};
	\node [label=right:{\small$\{(0,w,2)\}$}] at (3.8,3.9) {};
	\node [label=left:{\small$\{(w,m,2)\}$}] at (6.2,5) {};
	\node [label=right:{\small$\{(w,m,1)\}$}] at (-0.2,3.8) {};
	\node [label=right:{\small$\{(v,w,1)\}$}] at (1.8,4.1) {};
	\node [label=left:{\small$\{(0,v,1), (w,m,1)\}$}] at (-0.3,5.2) {};
	\node [label=right:{\small$\{(0,v,2)\}$}] at (2.8,6.4) {};
	\node [label=left:{\small$\{(w,m,1)\}$}] at (-1.3,7) {};
	\node [label=right:{\small$\{(0,v,1)\}$}] at (0.3,6.4) {};

@@

	% Numbers of extant ancestors and links, from top to bottom
	\node[label=right:{CA \; RE}] at (7.5, 8.2) {};
	\node[label=right:{$\binom{2}{2}$ \; $\frac{2 m - 2 w - 2}{m - 1} \rho$}] at (7.5, 7.6) {};
	\node[label=right:{$\binom{4}{2}$ \; $\frac{2 m + 2 v - 2 w - 4}{m - 1} \rho$}] at (7.5, 6.6) {};
	\node[label=right:{$\binom{3}{2}$ \; $\frac{2 m + v - w - 3}{m - 1} \rho$}] at (7.5, 5.7) {};
	\node[label=right:{$\binom{4}{2}$ \; $\frac{2 m + w - v - 4}{m - 1} \rho$}] at (7.5, 4.95) {};
	\node[label=right:{$\binom{5}{2}$ \; $\frac{2 m - 5}{m - 1} \rho$}] at (7.5, 3.8) {};
	\node[label=right:{$\binom{4}{2}$ \; $\frac{2 m - 4}{m -1} \rho$}] at (7.5, 2.75) {};
	\node[label=right:{$\binom{3}{2}$ \; $\frac{2 m - 3}{m - 1} \rho$}] at (7.5, 2.05) {};
	\node[label=right:{$\binom{2}{2}$ \; $2 \rho$}] at (7.5, 1.45) {};
	\node[label=right:{$\binom{3}{2}$ \; $3 \rho$}] at (7.5, 0.6) {};

@@

	\draw (0, 0) -- (0, 1.7) -- (-1,1.7) -- (1,1.7);
	\draw (-1.4, 1.5) -- (-1.1, 1.5) -- (-1.1, 1.4) -- (-1.4, 1.4) -- (-1.4, 1.5);
	\draw (-1.1, 1.45) -- (-0.6, 1.45);
	\draw (0.6, 1.45) -- (0.9, 1.45);
	\draw (0.9, 1.5) -- (1.4, 1.5) -- (1.4, 1.4) -- (0.9, 1.4) -- (0.9, 1.5);
	\draw (1.1, 1.6) -- (1.1, 1.3);
	\node [label=below:{\small $w$}] at (1.1, 1.5) {};

	\draw (5, 1.2) -- (5, 2.4) -- (4,2.4) -- (6,2.4);
	\draw (3.6, 2.2) -- (4.1, 2.2) -- (4.1, 2.1) -- (3.6, 2.1) -- (3.6, 2.2);
	\draw (4.1, 2.15) -- (4.4, 2.15);
	\draw (5.6, 2.15) -- (6.1, 2.15);
	\draw (6.1, 2.2) -- (6.4, 2.2) -- (6.4, 2.1) -- (6.1, 2.1) -- (6.1, 2.2);
	\draw [color=red](5.8, 2.3) -- (5.8, 2.0);
	\node [label={[red]below:{\small $y$}}] at (5.8, 2.1) {};

	\draw (1, 1.7) -- (1, 3.1) -- (0,3.1) -- (2,3.1);
	\draw (-0.4, 2.85) -- (0.1, 2.85);
	\draw (0.1, 2.9) -- (0.4, 2.9) -- (0.4, 2.8) -- (0.1, 2.8) -- (0.1, 2.9);
	\draw (1.6, 2.85) -- (1.9, 2.85);
	\draw (1.9, 2.9) -- (2.1, 2.9) -- (2.1, 2.8) -- (1.9, 2.8) -- (1.9, 2.9);
	\draw (2.1, 2.85) -- (2.4, 2.85);

	\draw (-1, 1.7) -- (-1, 4.5) -- (0,4.5) -- (0,3.1);
	\draw (-0.9, 4.3) -- (-0.6, 4.3) -- (-0.6, 4.2) -- (-0.9, 4.2) -- (-0.9, 4.3);
	\draw (-0.6, 4.25) -- (-0.4, 4.25);
	\draw (-0.4, 4.3) -- (-0.1, 4.3) -- (-0.1, 4.2) -- (-0.4, 4.2) -- (-0.4, 4.3);
	\draw (-0.5, 4.4) -- (-0.5, 4.1);
	\node [label=below:{\small $x$}] at (-0.5, 4.3) {};

@@

	% Gray dashed lines to visually separate holding times
	\draw[color=gray, dashed] (7.5, 1.2) -- (9.1, 1.2);
	\draw[color=gray, dashed] (7.5, 1.7) -- (9.1, 1.7);
	\draw[color=gray, dashed] (7.5, 2.4) -- (9.1, 2.4);
	\draw[color=gray, dashed] (7.5, 3.1) -- (9.1, 3.1);
	\draw[color=gray, dashed] (7.5, 3.8) -- (9.1, 3.8);
	\draw[color=gray, dashed] (7.5, 4.5) -- (9.1, 4.5);
	\draw[color=gray, dashed] (7.5, 5.4) -- (9.1, 5.4);
	\draw[color=gray, dashed] (7.5, 6) -- (9.1, 6);
	\draw[color=gray, dashed] (7.5, 6.6) -- (9.1, 6.6);
	\draw[color=gray, dashed] (7.5, 7.2) -- (9.1, 7.2);
\end{tikzpicture}
\end{tabular}
}
\caption{(A)
A realisation of the graph traversed by Hudson's algorithm started from a
sample of three chromosomes with $m$ discrete sites each at time $t = 0$, and
propagated until time $T$. The MRCA on the genetic interval $[v, w)$ is reached
at event \textsf{b}, while that on $[0, v)$ is reached at event \textsf{c}.
The non-ancestral segment $[v, w)$ above
\textsf{a} contributes to the rate of effective recombinations because it
is trapped between ancestral segments. The two columns titled CA and RE
are the respective rates of mergers and recombinations when
the recombination rate is $\rho$.
(B) A corresponding realisation of a big ARG, which augments Hudson's algorithm
by tracking nonancestral lineages. The result is a simpler state space and
dynamics, at the cost of extra nodes and edges, highlighted in red, which do
not affect the local tree at any site.
Recombination positions are labelled alphabetically in time, and their ordering
along the genome is $y < v < x < w$, of which the first only appears in panel B.
There are two separate recombination events at link $w$.
}
\label{hudson_vs_bigARG}
\end{figure}
